Tidy data structure to support exploration and modeling of temporal data

### 2018-12-29 main.Rmd
changed daily % plot aspect ratio
added tibbletime

Current time series objects, which are model-oriented matrices with little support for non-numeric values, feed into the framework solely at the modelling stage. Analysts are expected to do their own data pre-processing and take care of anything else needed to get to this stage, which leads to a myriad ad-hoc solutions and duplicated efforts. The diagram correspondingly would place the model at the centre of the analytical universe, and all the transformations, visualisations would hinge on that format.

@@

The **tidyverse** largely formalizes exploratory data analysis. The whole R community has quickly adopted the **tidyverse** way of thinking to much broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018]. Temporal data is left behind in this regard. This paper proposes a new tidy representation for temporal data and a unified framework to streamline the workflow from data pre-processing to visualization and forecasting, as an integral part of tidy data analysis.

@@

The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data. Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which we name as "tsibble", inspired by "ts" and "tibble". It bridges the gap between raw temporal data and model inputs. Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs. Index, key and time interval are the three stone pillars to this new semantically-structured temporal data.

@@

@@

Matrices are a conventional way of organizing time series, with standards being provided by native `ts` object in R, extended by **zoo** [@zoo2005], and **xts** [@R-xts]. Time indices are stored as an attribute, separating from the main body of the data.

@@

```{r sel-monthly-plot, fig.cap = "(ref:sel-monthly-plot-cap)", fig.height = 6}
```

>>>>

Current time series objects, which are model-oriented matrices with little support for non-numeric values, feed into the framework solely at the modelling stage. Analysts are expected to do their own data pre-processing and take care of anything else needed to get to this stage, which leads to a myriad ad-hoc solutions and duplicated efforts.

@@

The **tidyverse** largely formalizes exploratory data analysis. The whole R community has quickly adopted the **tidyverse** way of thinking to much broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018]. Temporal data is left behind in this regard.

## Existing time series standards in R

Current standards, provided by native `ts` object in R and extended by **zoo** [@zoo2005], and **xts** [@R-xts], organize temporal data in matrices with implicit time indices. These objects were designed for modelling methods. The diagram correspondingly would place the model at the center of the analytical universe, and all the transformations, visualisations would hinge on that format. This is contrary to the tidyverse conceptualisation, which wholistically captures the full data workflow.

A new temporal data class is needed in the upstream of the workflow, which could incorporate all the downstream modules. A relatively new R package **tibbletime** proposed a data class of *time tibble* to represent temporal data in heterogeneous tabular format. It only requires an index variable to declare a temporal data object, thus placing at the import stage. However, this is a loosely-defined structure and insufficient for time series analytics and models. Because it does not define interested subjects ($i$) over time and time uniqueness.

This paper proposes a new tidy representation for temporal data and a unified framework to streamline the workflow from data pre-processing to visualization and forecasting, as an integral part of tidy data analysis.

@@

The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data. Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which we refer to as "tsibble".

@@

It bridges the gap between raw temporal data and model inputs. Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs. Index, key and time interval are the three stone pillars to this new semantically-structured temporal data.

@@

@@

```{r sel-monthly-plot, fig.cap = "(ref:sel-monthly-plot-cap)", fig.height = 7}
```
